NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-06-26 paper/background.tex
grammar fix

\section{Background}
The target audiences of the proposed tool are domain experts who analyze and develop NLP models. Therefore, certain background knowledge in NLP is required to fully understand and appreciate the technique discussed in this paper. In this section, we first explain the definition of natural language inference (NLI) task and how it fits into the grand challenges in NLP. Then, we examine the common architectural characteristics shared by many state-of-the-art neural network models. Finally, we discuss the role attention plays in the model and why attention is closely tied to model interpretability.

@@

At first glance, the task of natural language inference may seem less practical compared to other well-known NLP challenges such as machine translation; however, the ability to distinguish entailment and contradiction relationship is fundamental to understanding natural language at large.
%
Considering the ambiguousness of natural language and polysemy of words, the inference task can become quite challenging (especially from the learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refers to premise, \textbf{H} refers to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal similarity between ``Electric'' and ``electrified'' may trick a model to predict \textbf{H2} as \emph{entailment}. A model likely will also fail to understand the link between ``went public'' and ``IPO'', and therefore, mistake \textbf{H1} as \emph{neutral}.
%
%The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
%deep comprehension of textual information. Major tasks include Natural Language Inference,
%Machine Translation, Questions Answering, Summarization, and etc.
%Recently, neural networks models have performed strongly on these tasks.
%
Recently, Bowman et al. introduced a large corpus~\cite{BowmanAngeliPotts2015} for NLI tasks, which has helped spawn a new wave of effective neural network models for those tasks. Here, we focus on the analysis of the decomposable attention model~\cite{parikh2016emnlp} on this dataset (see \textbf{Appendix A} for a detailed description of the model).

@@

These end-to-end models usually take pre-trained word vectors (numerical vector representation of individual words, in which the semantic similarities are expressed by distances in the high-dimensional vector space~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}) as input, and then in the encoding stage, the pre-trained vectors are adjusted to the specialized task at hand. Subsequently, the next stage aims at finding alignment between words in the input sequences. For the NLI task, this means finding the correspondence between words from the premise sentence to the hypothesis sentence (see details in Fig.~\ref{fig:attention} and Section~\ref{sec:attention}). Finally, in the last stage, the alignment information and the encoded vector representations are aggregated and then used as features for a classification network (last part of the end-to-end neural network).
%
Since all these three stages of the model are trained jointly in an end-to-end fashion, it is important to explore the interaction between intermediate representations and predictions to make sense of how predictions are made inside such a model.

\subsection{Attention Mechanism}
\label{sec:attention}
Among the three stages, the second stage often constitutes a crucial part of the model as it determines where the classifier will focus for generating a prediction. The operation to compute the alignment between words in the input is referred to as the attention mechanism~\cite{bahdanau2014neural}. 
The introduction of the attention mechanism allows
pairwise interaction between internal representations of words. 
This interaction can be naturally explained as a form of alignment that exposes an interpretable layer in end-to-end neural networks.
%
%As the only interpretable part of the internal state of the network, understanding how attention information 
Recently, attention has contributed to the strong performance of many NLP models~\cite{parikh2016emnlp,rush2015neural,yang2016hierarchical,seo2016bidirectional,schwartz2017high}.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.85\linewidth]{attentionIllustration}
  \vspace{-2mm}
 \caption{Attention can be naturally explained as a form of alignment which exposes an interpretable layer in end-to-end neural networks. On the left, the matrix shows the attention between the premise and hypothesis pair. On the right, we illustrate that the attention can also be interpreted as alignment.}
\label{fig:attention}
\vspace{-2mm}
\end{figure}

For natural language inference task, as illustrated in Fig.~\ref{fig:attention}, the attention information can be represented as a matrix describing the soft alignment between words in the premise and the words in the hypothesis. As we can see in this example, the higher values indicate significant alignments between words. The \emph{subject}, \emph{verb}, and \emph{object} between these two sentences are aligned correctly (words, such as ``an'', which has less importance in determining sentence relationship, was assigned a lower score). Interestingly, the difference between the subject words does not affect their alignment. The classification stage can then utilize this information to determine the subject of the sentence is different, and therefore, despite the other part of the sentences are identical, the relationship between the premise and the hypothesis is neutral.

>>>>

\section{Background}
The target audience of the proposed tool is domain experts who analyze and develop NLP models. Therefore, certain background knowledge in NLP is required to fully understand and appreciate the technique discussed in this paper. In this section, we first explain the definition of a natural language inference (NLI) task and how it fits into the grand challenges in NLP. Then, we examine the common architectural characteristics shared by many state-of-the-art neural network models. Finally, we discuss the role attention plays in the model and why attention is closely tied to model interpretability.

@@

At first glance, the task of natural language inference may seem less practical compared to other well-known NLP challenges such as machine translation; however, the ability to distinguish the entailment and contradiction relationship is fundamental to understanding natural language at large.
%
Considering the ambiguousness of natural language and the polysemy of words, the inference task can become quite challenging (especially from the learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refers to premise, \textbf{H} refers to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal similarity between ``Electric'' and ``electrified'' may trick a model to predict \textbf{H2} as \emph{entailment}. A model likely will also fail to understand the link between ``went public'' and ``IPO'', and therefore, mistake \textbf{H1} as \emph{neutral}.
%
%The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
%deep comprehension of textual information. Major tasks include Natural Language Inference,
%Machine Translation, Questions Answering, Summarization, and etc.
%Recently, neural networks models have performed strongly on these tasks.
%
Recently, Bowman et al. introduced a large corpus~\cite{BowmanAngeliPotts2015} for NLI tasks, which has helped spawn a new wave of effective neural network models for those tasks. Here, we focus on the analysis of the decomposable attention model~\cite{parikh2016emnlp} on this dataset (see \textbf{Appendix A} for a detailed description of the model).

@@

These end-to-end models usually take pre-trained word vectors (numerical vector representation of individual words, in which the semantic similarities are expressed by distances in the high-dimensional vector space~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}) as input, and then in the encoding stage, the pre-trained vectors are adjusted to the specialized task at hand. Subsequently, the next stage aims at finding alignment between words in the input sequences. For the NLI task, this means finding the correspondence between words from the premise sentence to the hypothesis sentence (see details in Fig.~\ref{fig:attention} and Section~\ref{sec:attention}). Finally, in the last stage, the alignment information and the encoded vector representations are aggregated and then used as features for a classification network (last part of the end-to-end neural network).
%
Since all three stages of the model are trained jointly in an end-to-end fashion, it is important to explore the interaction between intermediate representations and predictions to make sense of how predictions are made inside such a model.

\subsection{Attention Mechanism}
\label{sec:attention}
Among the three stages, the second stage often constitutes a crucial part of the model as it determines where the classifier will focus for generating a prediction. The operation to compute the alignment between words in the input is referred to as the attention mechanism~\cite{bahdanau2014neural}. 
The introduction of the attention mechanism allows
pairwise interaction between internal representations of words. 
This interaction can be naturally explained as a form of alignment that exposes an interpretable layer in end-to-end neural networks.
%
%As the only interpretable part of the internal state of the network, understanding how attention information 
Recently, the attention mechanism has contributed to the strong performance of many NLP models~\cite{parikh2016emnlp,rush2015neural,yang2016hierarchical,seo2016bidirectional,schwartz2017high}.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.85\linewidth]{attentionIllustration}
  \vspace{-2mm}
 \caption{Attention can be naturally explained as a form of alignment that exposes an interpretable layer in end-to-end neural networks. On the left, the matrix shows the attention between the premise and hypothesis pair. On the right, we illustrate that the attention can also be interpreted as alignment.}
\label{fig:attention}
\vspace{-2mm}
\end{figure}

For natural language inference task, as illustrated in Fig.~\ref{fig:attention}, the attention information can be represented as a matrix describing the soft alignment between words in the premise and words in the hypothesis. As we can see in this example, the higher values indicate significant alignments between words. The \emph{subject}, \emph{verb}, and \emph{object} between these two sentences are aligned correctly (words, such as ``an'', which has less importance in determining sentence relationship, was assigned a lower score). Interestingly, the difference between the subject words does not affect their alignment. The classification stage can then utilize this information to determine if the subject of the sentence is different, and therefore, despite the other part of the sentences being identical, the relationship between the premise and the hypothesis is neutral.
